A numerical study of the solution of X-Mode equations around the hybrid resonance

### 2015-05-04 report_clean/responce_reviewers.tex

\end{abstract}

\section{General}

We marked corrections suggested by the reviewer 1 in blue, by the reviewer 2 in red, 
and some extra corrections made by us that can be important in green.

More important corrections include

- changed the wording in the abstract

- corrected the definition of the cold plasma dielectric tensor, to match the convention $\partial_t\rightarrow -i\omega$ (it had to be hermitian conjugate; 
this had been done everywhere, the corresponding corrections had been made in Table 1 and some figures)

- corrected the definitions of $\alpha,\delta$ to comply with the definition of the cold plasma dielectric tensor of Stix

(more precisely, removed terms $\frac{\omega^2}{c^2}$ in 
front of them)

- corrected the formulation of the lemma on the well-posedness and added a remark after the lemma on the sign of $\nu$.

-corrected assumption on smoothness of $E_y^{\nu}$ in the section 'Numerical Experiments'.

\section{Response to the reviewer 1}

@@

- p12 Section 4.3.1

  It is true, but in this case there is no resonance, since $\delta(x)\equiv 0$. Hence $E_x=-\frac{i\delta}{\alpha}E_y\equiv 0$, 
  and $E_y$ solves the Airy equation in the frequency domain. We added a small to Section 4.3.1 to underline this.

- p12 Equation (23)
We changed it to $e$

- p14 Figures 8 and 9
Added the captions

- p15 in Conclusion
  Indeed $P1$!

 -references: corrected

### 2015-05-04 report_clean/absorption.tex
plus a few small bugs fixed

In all the experiments in this section the CFL number was chosen as 0.5.
\subsubsection{No-Resonance Case}
\label{section:absorption_airy}
We choose the parameters so that in the frequency domain, for the limiting amplitude problem, $\hat{E}_{y}^{\nu},\; \nu=0,$ satisfies 
the Airy equation, c.f. also Remark \ref{remark:other}. We set $\omega_c=0$ (thus $\delta(x)=0$), $\omega=1$ (hence $\alpha(x)=1-N_e(x)$), 
choose the domain as $(-0.5, 10)$ and set the electron density $N_e(x)=1+x>0,\; x\in \Omega$. 
\urev{Importantly, since $\delta(x)=0$, in this case no resonance occurs and $\hat{E}_x(x)\equiv 0$. This corresponds to the case when the
tensor of dielectric permittivity $\underline{\varepsilon}^{\nu}$ is 
diagonal.}

The boundary conditions in (\ref{eq:bcs}) are chosen as $G=Ai'(0.5)$.

>>>>

In all the experiments in this section the CFL number was chosen as 0.5.
\subsubsection{No-Resonance Case}
\label{section:absorption_airy}
We choose the parameters so that in the frequency domain, for the limiting amplitude problem, $\hat{E}_{y}^{\nu},\; \nu=0,$ satisfies 
the Airy equation, c.f. also Remark \ref{remark:other}. We set $\omega_c=0$ (thus $\delta(x)=0$), $\omega=1$ (hence $\alpha(x)=1-N_e(x)$), 
choose the domain as $(-0.5, 10)$ and set the electron density $N_e(x)=1+x>0,\; x\in \Omega$. 
\urev{Importantly, since $\delta(x)=0$, in this case no resonance occurs and $\hat{E}_x(x)\equiv 0$. This corresponds to the case when the
tensor of dielectric permittivity $\uuline{\varepsilon^{\nu}_{\omega}}$ is 
diagonal.}

The boundary conditions in (\ref{eq:bcs}) are chosen as $G=Ai'(-0.5)$.

### 2015-05-04 report_clean/intro.tex
plus a few small bugs fixed

To illustrate unexpected difficulties that can occur in this system, let us rewrite it in the frequency domain. To do so, 
we introduce the plasma frequency $\omega_p(\mathbf{r})=\sqrt{\frac{|e|^2N_e(\mathbf{r})}{m\epsilon_0}}$, 
the cyclotron frequency $\omega_c(\mathbf{r})=\frac{e B_0(\mathbf{r})}{m_e}$ and perform the Fourier transform 
in time of (\ref{eq:main_system}) (with the convention $\partial_t\rightarrow -i\omega$), for a 
particular case $\nu=0$:
\begin{align}
\label{eq:model_freq_domain}
 \mathbf{curl}{\operatorname{curl}}\hat{\mathbf{E}}-\frac{\omega^2}{c^2}\uuline{\varepsilon_{\omega}}^{0}(\mathbf{r})\hat{\mathbf{E}}=0,
\end{align}
where the cold-plasma dielectric tensor \cite[Chapter 1-2]{Stix}
\begin{align}
\nonumber
\uuline{\varepsilon_{\omega}}^{0}(\mathbf{r})=\left(
 \begin{matrix}
  \alpha_{\omega}(\mathbf{r})& -i\delta_{\omega}(\mathbf{r})\\
  i\delta_{\omega}(\mathbf{r}) & \alpha_{\omega}(\mathbf{r})
 \end{matrix}
\right),\\
\label{eq:alpha_delta}
\alpha_{\omega}(\mathbf{r})=1-\frac{\omega_p^2(\mathbf{r})}{\omega^2-\omega_c^2(\mathbf{r})},\; \delta_{\omega}(\mathbf{r})=\frac{\omega_c(\mathbf{r})\omega_p^2(\mathbf{r})}{\omega(\omega^2-\omega_c^2(\mathbf{r}))}.
\end{align}

The point $\mathbf{r}_0$ s.t. $\omega_c(\mathbf{r}_0)=\omega$, i.e. when $\uuline{\varepsilon_{\omega}}^{0}(\mathbf{r}_0)$ is singular, is called a cyclotron resonance (c.f. \cite[Chapter 1-5]{Stix}). 
In \cite{singular_solutions} it was demonstrated for a 1D-counterpart of the above system that such points behave like removable singularities, 
hence this case is not of interest for the present work.

Another interesting case is $\omega_p^2(\mathbf{r})+\omega_c^2(\mathbf{r})=\omega^2$ and $\omega_c(\mathbf{r})\omega_p^2(\mathbf{r})\neq 0$, i.e. when the diagonal part of the tensor 
$\uuline{\varepsilon_{\omega}}^{0}(\mathbf{r})$ vanishes and the non-diagonal part remains non-zero. This is a hybrid resonance case.

@@

In the frequency domain (\ref{eq:main_model}) for $\nu=0$ is of the form 
\begin{align}
\label{eq:main_frequency_domain_intro}
\left(
\begin{matrix}
 i\theta \partial_x \hat{E}_y+\theta^2 \hat{E}_x\\
 i\theta \partial_x \hat{E}_x -\partial_x^2 \hat{E}_y
\end{matrix}
\right)-\frac{\omega^2}{c^2}
\uuline{\varepsilon_{\omega}}^{0}(x)\left(
\begin{matrix}
 \hat{E}_x\\
 \hat{E}_y
\end{matrix}
\right)
=0,
\end{align}
see (\ref{eq:alpha_delta}). Indeed, the coefficients $\alpha_{\omega},\;\delta_{\omega}$ depend solely on $x$ and $\omega$.  
\arev{
We concentrate on a physical situation of the hybrid resonance, which we define as in \cite{singular_solutions}.
\begin{df}
Let $\omega\in\mathbb{R}$ be fixed. 
We will call a point $x_h\in \mathbb{R}$ a hybrid resonance if $\alpha_{\omega}$ has 
a simple root in $x_h$ and has no other roots in an $\epsilon$-vicinity of this point $B_{\epsilon}(x_h)$, for some $\epsilon>0$, and 
$\delta(x_h)\neq 0$. 
\end{df}
}

To investigate the behaviour of the solutions in the case when the hybrid resonance occurs in the sustem, 
instead of the above problem we will study a sequence of the regularized problems, for $\nu\rightarrow 0$, 
\begin{align}
\label{eq:seq_regularized}
\left(
\begin{matrix}
 i\theta \partial_x \hat{E}_y+\theta^2 \hat{E}_x\\
 i\theta \partial_x \hat{E}_x -\partial_x^2 \hat{E}_y
\end{matrix}
\right)-\frac{\omega^2}{c^2}
\left(\uuline{\varepsilon_{\omega}}{0}(x)+i\nu \mathrm{Id}\right)\left(
\begin{matrix}
 \hat{E}_x\\
 \hat{E}_y
\end{matrix}
\right)
=0,
\end{align}
where $\mathrm{Id}$ is the identity matrix. More details on the connection of this problem with (\ref{eq:main_model}) will be provided in further sections.

>>>>

To illustrate unexpected difficulties that can occur in this system, let us rewrite it in the frequency domain. To do so, 
we introduce the plasma frequency $\omega_p(\mathbf{r})=\sqrt{\frac{|e|^2N_e(\mathbf{r})}{m\epsilon_0}}$, 
the cyclotron frequency $\omega_c(\mathbf{r})=\frac{e B_0(\mathbf{r})}{m_e}$ and perform the Fourier transform 
in time of (\ref{eq:main_system}) (with the convention $\partial_t\rightarrow -i\omega$), for a 
particular case $\nu=0$:
\begin{align}
\label{eq:model_freq_domain}
 \mathbf{curl}{\operatorname{curl}}\hat{\mathbf{E}}-\frac{\omega^2}{c^2}\uuline{\varepsilon_{\omega}^{0}}(\mathbf{r})\hat{\mathbf{E}}=0,
\end{align}
where the cold-plasma dielectric tensor \cite[Chapter 1-2]{Stix}
\begin{align}
\nonumber
\uuline{\varepsilon_{\omega}^{0}}(\mathbf{r})=\left(
 \begin{matrix}
  \alpha_{\omega}(\mathbf{r})& -i\delta_{\omega}(\mathbf{r})\\
  i\delta_{\omega}(\mathbf{r}) & \alpha_{\omega}(\mathbf{r})
 \end{matrix}
\right),\\
\label{eq:alpha_delta}
\alpha_{\omega}(\mathbf{r})=1-\frac{\omega_p^2(\mathbf{r})}{\omega^2-\omega_c^2(\mathbf{r})},\; \delta_{\omega}(\mathbf{r})=\frac{\omega_c(\mathbf{r})\omega_p^2(\mathbf{r})}{\omega(\omega^2-\omega_c^2(\mathbf{r}))}.
\end{align}

The point $\mathbf{r}_0$ s.t. $\omega_c(\mathbf{r}_0)=\omega$, i.e. when $\uuline{\varepsilon_{\omega}^{0}}(\mathbf{r}_0)$ is singular, is called a cyclotron resonance (c.f. \cite[Chapter 1-5]{Stix}). 
In \cite{singular_solutions} it was demonstrated for a 1D-counterpart of the above system that such points behave like removable singularities, 
hence this case is not of interest for the present work.

Another interesting case is $\omega_p^2(\mathbf{r})+\omega_c^2(\mathbf{r})=\omega^2$ and $\omega_c(\mathbf{r})\omega_p^2(\mathbf{r})\neq 0$, i.e. when the diagonal part of the tensor 
$\uuline{\varepsilon_{\omega}^{0}}(\mathbf{r})$ vanishes and the non-diagonal part remains non-zero. This is a hybrid resonance case.

@@

In the frequency domain (\ref{eq:main_model}) for $\nu=0$ is of the form 
\begin{align}
\label{eq:main_frequency_domain_intro}
\left(
\begin{matrix}
 i\theta \partial_x \hat{E}_y+\theta^2 \hat{E}_x\\
 i\theta \partial_x \hat{E}_x -\partial_x^2 \hat{E}_y
\end{matrix}
\right)-\frac{\omega^2}{c^2}
\uuline{\varepsilon_{\omega}^{0}}(x)\left(
\begin{matrix}
 \hat{E}_x\\
 \hat{E}_y
\end{matrix}
\right)
=0,
\end{align}
see (\ref{eq:alpha_delta}). Indeed, the coefficients $\alpha_{\omega},\;\delta_{\omega}$ depend solely on $x$ and $\omega$.  
\arev{
We concentrate on a physical situation of the hybrid resonance, which we define as in \cite{singular_solutions}.
\begin{df}
Let $\omega\in\mathbb{R}$ be fixed. 
We will call a point $x_h\in \mathbb{R}$ a hybrid resonance if $\alpha_{\omega}$ has 
a simple root in $x_h$ and has no other roots in an $\epsilon$-vicinity of this point $B_{\epsilon}(x_h)$, for some $\epsilon>0$, and 
$\delta(x_h)\neq 0$. 
\end{df}
}

To investigate the behaviour of the solutions in the case when the hybrid resonance occurs in the sustem, 
instead of the above problem we will study a sequence of the regularized problems, for $\nu\rightarrow 0$, 
\begin{align}
\label{eq:seq_regularized}
\left(
\begin{matrix}
 i\theta \partial_x \hat{E}_y+\theta^2 \hat{E}_x\\
 i\theta \partial_x \hat{E}_x -\partial_x^2 \hat{E}_y
\end{matrix}
\right)-\frac{\omega^2}{c^2}
\left(\uuline{\varepsilon_{\omega}^{0}}(x)+i\nu \mathrm{Id}\right)\left(
\begin{matrix}
 \hat{E}_x\\
 \hat{E}_y
\end{matrix}
\right)
=0,
\end{align}
where $\mathrm{Id}$ is the identity matrix. More details on the connection of this problem with (\ref{eq:main_model}) will be provided in further sections.
